---
title: "WSW Study 3: How many participants do we need to have enough power to find a US valence x Task interaction effect on evaluative ratings?"
author: "Jérémy Béna"
date: today
format: 
  html:
    toc: true
    published-title: "Date"
    code-fold: true
    code-summary: "Display R code"
    code-tools: true
embed-resources: true
execute: 
  warning: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, comment = NA)
```

```{r read_data}
#Load (or install and load) packages
if (!require("pacman")) install.packages("pacman")
p_load('jsonlite', 'tidyverse', 'psych', 'papaja', 'kableExtra', 'afex', 'emmeans', 'effectsize', 'Superpower', 'gt', 'ggstatsplot') 

set_sum_contrasts()

#read data files
read_file('data/pilot2_age.txt')%>%
#... split it into lines ...
 str_split('\n') %>% first() %>%
#... filter empty rows ...
 discard(function(x) x == '') %>%
#... parse JSON into a data.frame
 map_dfr(fromJSON, flatten=T) -> dat_age

read_file('data/pilot2_val.txt')%>%
#... split it into lines ...
 str_split('\n') %>% first() %>%
#... filter empty rows ...
 discard(function(x) x == '') %>%
#... parse JSON into a data.frame
 map_dfr(fromJSON, flatten=T) -> dat_val

#arrange
dat = full_join(dat_age, dat_val)

dat$url.srid = as.factor(dat$url.srid)
dat$us_valence = as.factor(dat$us_valence)
dat$reco_resp = as.factor(dat$reco_resp)

dat = dat %>%
   fill(url.srid, .direction = "down") %>% 
  group_by(url.srid) %>% 
  fill(consent, comment_study,pay_attention,serious, instructions_conditioning, '-0','-1','-2','-3','-4','-5','-6', .direction = "down") %>%
  fill(consent, comment_study, pay_attention,serious,instructions_conditioning,'-0','-1','-2','-3','-4','-5','-6', .direction = "up") %>%
  select(url.srid, sender, consent, duration, ended_on, pay_attention, serious, response, response_action 
      #general information
         , '-0','-1','-2','-3','-4','-5','-6',comment_study,count_trial_learning, cs, us, us_valence, us_age, uss, resp_pos_learning #learning phase
         , count_trial_memory, idtarg, reco_resp, source_mem  #memory phase
         , count_trial_ratings, cs_rating #rating phase
         , instructions_conditioning 
          ) 

dat$sports <- ifelse(dat$'-1'==TRUE|
                       dat$'-2'==TRUE|
                       dat$'-3'==TRUE|
                       dat$'-4'==TRUE|
                       dat$'-5'==TRUE|
                       dat$'-0'==TRUE|
                       dat$'-6'==TRUE,0,1)

#exclude
dat = dat %>% filter(sports == 1 & pay_attention == 1 & serious == 1)
```

# Added scenario in version 2 (18 February)

Update on 18 February: added Scenario 2b, which is Scenario 2 with $\alpha$ = .10 (instead of .05) and $1 - \beta$ = .9 (instead of .8). In this scenario, we would need _n_ = 52 in each learning instructions condition. 

# This document 

I first present the target 2 (US valence) $\times$ 2 (Learning instructions) mixed ANOVA on mean evaluative ratings and correlations between repeated measures within each Learning instructions condition. I then present power analyses in four scenarios to plan for sample size to detect (1) a US valence $\times$ Learning instructions interaction (larger EC effect in the valence focus than in the age focus condition) and (2) a simple effect of US valence in the valence focus condition. Note that these estimates may be conservative, as I used an $\alpha$ level of .05 in all analyses (I did not adapt the $\alpha$ level for one-sided tests). 

# Summary

If the interaction is exactly the one found in Pilot study 2 (with identical means, sds, and correlations; Scenario 1), we would need a total of 50 participants. If we have reasons to suspect a smaller interaction (Scenarios 2 and 3), we would need up to a total of 264 participants (Scenario 3; Scenario 2 is less pessimistic with a total of 96 participants). If we expect smaller (but still positive) correlations between mean evaluative ratings within each Learning instructions condition, we may need too many participants, too (Scenario 4). 

# Calculate means and sds in each US Valence $\times$ Learning instructions cell

```{r ratings}
ratings = subset(dat,sender == "rating_trial")

ratings$cs_rating = as.numeric(ratings$cs_rating)
ratings$us_age[is.na(ratings$us_age)] = "dist"
ratings = ratings %>% mutate(cs_sex = ifelse(grepl("m_m", cs), "male", "female")) 

ratings_cs = ratings %>% filter(us_valence!="dist") %>% droplevels()

rat = ratings_cs %>% group_by(url.srid, us_valence, instructions_conditioning) %>%
  summarize(mean_rating = mean(cs_rating))

rat_agg = describeBy(rat$mean_rating, list(rat$us_valence, rat$instructions_conditioning), mat = TRUE)

gt(describeBy(rat$mean_rating, list(rat$us_valence, rat$instructions_conditioning), mat = TRUE, digits = 2), caption = "Descriptive statistics of mean evaluative ratings as a function of US valence ('group1') and Learning instructions ('group2')")
```

# Check the interaction in a $2 \times 2$ mixed ANOVA on mean evaluative ratings

```{r ec_effect}
mod1 = aov_ez(data=rat
       ,id="url.srid"
       ,dv="mean_rating"
       ,within ="us_valence"
       ,between="instructions_conditioning"
       )

p_mod1 = apa_print(mod1, es = "pes")

s_eff = pairs(emmeans(mod1,~us_valence|instructions_conditioning))

p_s_eff = apa_print(s_eff)
```

The US valence $\times$ Learning instructions interaction was significant, `r p_mod1$full_result$instructions_conditioning_us_valence`. The EC effect (neg-pos) was significant in the Valence focus condition, `r p_s_eff$full_result$Val_task_Negative_positive` but not in the Age focus condition, `r p_s_eff$full_result$Age_task_Negative_positive` (see Fig. below).

```{r figure, fig.cap="Mean evaluative ratings as a function of US valence and Learning instructions"}
grouped_ggwithinstats(data = rat
    , x = us_valence
    , grouping.var = instructions_conditioning
    , y = mean_rating,
    , xlab = "US Valence",
    , ylab = "Mean evaluative ratings"
    , results.subtitle = FALSE
)
```

# Correlations between mean evaluative ratings for CS paired with US+ vs. US- in each Learning instructions condition

```{r cor}
rat_age = rat %>% filter(instructions_conditioning=="age_task")
p_ca = apa_print(cor.test(rat_age$mean_rating[rat_age$us_valence=="positive"], rat_age$mean_rating[rat_age$us_valence=="negative"]))

rat_val = rat %>% filter(instructions_conditioning=="val_task")
p_cv = apa_print(cor.test(rat_val$mean_rating[rat_val$us_valence=="positive"], rat_val$mean_rating[rat_val$us_valence=="negative"]))
```

Mean evaluative ratings for CS paired with US+ and US- were positively associated in each Learning instructions condition (valence focus: `r p_cv$estimate`; age focus: `r p_ca$estimate`).

# Four scenarios for the power analyses: From quite optimistic to less so

Find required _n_ (per Learning instructions condition) to have enough statistical power ($1-\beta = .8$) to detect the US valence $\times$ Learning instructions interaction on mean evaluative ratings (+ the EC effect in the valence focus condition) in four scenarios:

- **Scenario 1:** Exact means, SDs, and correlations as in Pilot Study 2
- **Scenario 2:** Scenario 1 but increase (reduce) the EC effect in the age (valence) focus condition to reduce the interaction effect (+ scenario 2b, with $\alpha$ = .1 and $1 - \beta$ = .9)
- **Scenario 3:** Scenario 2 but reduce the interaction effect even more 
- **Scenario 4:** Scenario 2 but reduce correlations to $r = .5$

Additional scenarios (negative and null correlations between repeated measures) are considered in the "Additional analyses" section.

Analyses were conducted with at least $n =$ 10 and maximum $n =$ 150 in each Learning instructions condition.

## Scenario 1: Exact means, SDs, and correlations as in Pilot Study 2

```{r setup_design1}
n = 50 #just to generate the anova design

r_age = cor(rat_age$mean_rating[rat_age$us_valence=="positive"],rat_age$mean_rating[rat_age$us_valence=="negative"]) #correlation between ratings on CS paired with US+ and US- in the age condition

r_val = cor(rat_val$mean_rating[rat_val$us_valence=="positive"], rat_val$mean_rating[rat_val$us_valence=="negative"]) #correlation between ratings on CS paired with US+ and US- in the valence condition

r_mat = matrix(c(1,r_age,0,0
         ,r_age,1,0,0
         ,0,0,1,r_val
         ,0,0,r_val,1), nrow=4) #use values from the correlations calculated above 

design_exp = ANOVA_design(design = "2b*2w"
                          ,label_list = c("learn_instr"
                                          , "age", "valence"
                                          ,"us_val"
                                          ,"neg", "pos")
                           ,n = n
                           ,mu = c(rat_agg$mean[1], rat_agg$mean[2], rat_agg$mean[3], rat_agg$mean[4]) 
                           ,sd = c(rat_agg$sd[1], rat_agg$sd[2], rat_agg$sd[3], rat_agg$sd[4])
                           ,r=r_mat
                           ,plot = TRUE)
```

```{r results_power1}
n_min = 10
n_max = 150
pwr = .8
a_lev = .05

#Find statistical power for a range of sample sizes
power_design1 = plot_power(design_exp
                          ,min_n = n_min
                          ,max_n = n_max
                          ,desired_power = pwr*100
                          ,emm = TRUE
                          ,emm_comp = "us_val|learn_instr"
                          ,alpha_level = a_lev
                          ,plot = FALSE
                          ,verbose = FALSE) 

#power_design1$plot_emm

gt(power_design1$anova_n,caption = "Power analysis for the ANOVA -- Scenario 1")

gt(power_design1$effect_sizes,caption = "Effect sizes in the ANOVA -- Scenario 1")

gt(power_design1$emm_n,caption = "Power analysis for the post hoc analyses -- Scenario 1")

gt(power_design1$effect_sizes_emm,caption = "Effect sizes of the EC effects -- Scenario 1")
```

In **Scenario 1**, we would need at least **`r power_design1$anova_n$n[3]` participants in each Learning instructions condition** ($N_{total}$ = `r power_design1$anova_n$n[3]*2`) to have enough power ($1-\beta$ = `r power_design1$anova_n$desired_power[3]/100`; calculated: `r round(power_design1$anova_n$achieved_power[3]/100, digits=2)`) to find the interaction (estimated to be $\eta^2_p$ $\approx$ `r round(power_design1$effect_sizes$partial_eta_squared[3], digits=3)`). 

This sample size provides enough power to find an EC effect of $\eta^2_p$ $\approx$ `r round(power_design1$effect_sizes_emm$partial_eta_squared[2], digits=3)` in the **Valence** focus condition, as $1 - \beta$ = `r round(power_design1$emm_n$achieved_power[2]/100, digits=2)` to find this effect is achieved with $n$ = `r power_design1$emm_n$n[2]`. 

Regarding the EC effect in the **Age** focus condition, $n$ = `r power_design1$emm_n$n[1]` `r ifelse(power_design1$emm_n$n[1]==100, "(maximum n in the analysis)", "")` is `r ifelse(power_design1$emm_n$achieved_power[1]<80, "not enough", "needed")` to detect it ($1-\beta$ = `r round(power_design1$emm_n$achieved_power[1]/100, digits=2)`; $\eta^2_p$ $\approx$ `r round(power_design1$effect_sizes_emm$partial_eta_squared[1], digits=3)`).

## Scenario 2: Scenario 1 but reduce the interaction effect

I tried to reduce the interaction term without affecting much main effects. I increased the EC effect in the age focus condition and reduced it in the valence focus condition to calculate power to detect both a smaller interaction and a smaller EC effect in the valence focus condition.

```{r setup_design2}
n = 50 #just to generate the anova design

r_age = cor(rat_age$mean_rating[rat_age$us_valence=="positive"],rat_age$mean_rating[rat_age$us_valence=="negative"]) #correlation between ratings on CS paired with US+ and US- in the age condition

r_val = cor(rat_val$mean_rating[rat_val$us_valence=="positive"], rat_val$mean_rating[rat_val$us_valence=="negative"]) #correlation between ratings on CS paired with US+ and US- in the valence condition

reduce_inter = .04 #changes EC effects by .08; reduces interaction by .16

r_mat = matrix(c(1,r_age,0,0
         ,r_age,1,0,0
         ,0,0,1,r_val
         ,0,0,r_val,1), nrow=4) #use values from the correlations calculated above 

design_exp2 = ANOVA_design(design = "2b*2w"
                          ,label_list = c("learn_instr"
                                          , "age", "valence"
                                          ,"us_val"
                                          ,"neg", "pos")
                           ,n = n
                           ,mu = c(rat_agg$mean[1]-reduce_inter, rat_agg$mean[2]+reduce_inter, rat_agg$mean[3]+reduce_inter, rat_agg$mean[4]-reduce_inter) 
                           ,sd = c(rat_agg$sd[1], rat_agg$sd[2], rat_agg$sd[3], rat_agg$sd[4])
                           ,r=r_mat
                           ,plot = TRUE)
```

```{r results_power2}
n_min = 10
n_max = 150
pwr = .8
a_lev = .05

#Find statistical power for a range of sample sizes
power_design2 = plot_power(design_exp2
                          ,min_n = n_min
                          ,max_n = n_max
                          ,desired_power = pwr*100
                          ,emm = TRUE
                          ,emm_comp = "us_val|learn_instr"
                          ,alpha_level = a_lev
                          ,plot = FALSE
                          ,verbose = FALSE) 

#power_design2$plot_emm

gt(power_design2$anova_n,caption = "Power analysis for the ANOVA -- Scenario 2")

gt(power_design2$effect_sizes,caption = "Effect sizes in the ANOVA -- Scenario 2")

gt(power_design2$emm_n,caption = "Power analysis for the post hoc analyses -- Scenario 2")

gt(power_design2$effect_sizes_emm,caption = "Effect sizes of the EC effects -- Scenario 2")
```

In **Scenario 2**, we would need at least **`r power_design2$anova_n$n[3]` participants in each Learning instructions condition** ($N_{total}$ = `r power_design2$anova_n$n[3]*2`) to have enough power ($1-\beta$ = `r power_design2$anova_n$desired_power[3]/100`; calculated: `r round(power_design2$anova_n$achieved_power[3]/100, digits=2)`) to find the interaction (estimated to be $\eta^2_p$ $\approx$ `r round(power_design2$effect_sizes$partial_eta_squared[3], digits=3)`). 

This sample size provides enough power to find an EC effect of $\eta^2_p$ $\approx$ `r round(power_design2$effect_sizes_emm$partial_eta_squared[2], digits=3)` in the **Valence** focus condition, as $1 - \beta$ = `r round(power_design2$emm_n$achieved_power[2]/100, digits=2)` to find this effect is achieved with $n$ = `r power_design2$emm_n$n[2]`. 

Regarding the EC effect in the **Age** focus condition, $n$ = `r power_design2$emm_n$n[1]` `r ifelse(power_design2$emm_n$n[1]==100, "(maximum n in the analysis)", "")` is `r ifelse(power_design2$emm_n$achieved_power[1]<80, "not enough", "needed")` to detect it ($1-\beta$ = `r round(power_design2$emm_n$achieved_power[1]/100, digits=2)`; $\eta^2_p$ $\approx$ `r round(power_design2$effect_sizes_emm$partial_eta_squared[1], digits=3)`).

## Scenario 2b: Scenario 2 but alpha level = .1 and power = .9

```{r setup_design2b}
n = 50 #just to generate the anova design

r_age = cor(rat_age$mean_rating[rat_age$us_valence=="positive"],rat_age$mean_rating[rat_age$us_valence=="negative"]) #correlation between ratings on CS paired with US+ and US- in the age condition

r_val = cor(rat_val$mean_rating[rat_val$us_valence=="positive"], rat_val$mean_rating[rat_val$us_valence=="negative"]) #correlation between ratings on CS paired with US+ and US- in the valence condition

reduce_inter = .04 #changes EC effects by .08; reduces interaction by .16

r_mat = matrix(c(1,r_age,0,0
         ,r_age,1,0,0
         ,0,0,1,r_val
         ,0,0,r_val,1), nrow=4) #use values from the correlations calculated above 

design_exp2b = ANOVA_design(design = "2b*2w"
                          ,label_list = c("learn_instr"
                                          , "age", "valence"
                                          ,"us_val"
                                          ,"neg", "pos")
                           ,n = n
                           ,mu = c(rat_agg$mean[1]-reduce_inter, rat_agg$mean[2]+reduce_inter, rat_agg$mean[3]+reduce_inter, rat_agg$mean[4]-reduce_inter) 
                           ,sd = c(rat_agg$sd[1], rat_agg$sd[2], rat_agg$sd[3], rat_agg$sd[4])
                           ,r=r_mat
                           ,plot = TRUE)
```

```{r results_power2b}
n_min = 10
n_max = 150
pwr = .9
a_lev = .1

#Find statistical power for a range of sample sizes
power_design2b = plot_power(design_exp2b
                          ,min_n = n_min
                          ,max_n = n_max
                          ,desired_power = pwr*100
                          ,emm = TRUE
                          ,emm_comp = "us_val|learn_instr"
                          ,alpha_level = a_lev
                          ,plot = FALSE
                          ,verbose = FALSE) 

#power_design2$plot_emm

gt(power_design2b$anova_n,caption = "Power analysis for the ANOVA -- Scenario 2")

gt(power_design2b$effect_sizes,caption = "Effect sizes in the ANOVA -- Scenario 2")

gt(power_design2b$emm_n,caption = "Power analysis for the post hoc analyses -- Scenario 2")

gt(power_design2b$effect_sizes_emm,caption = "Effect sizes of the EC effects -- Scenario 2")
```

In **Scenario 2b** (which is Scenario 2 but with $\alpha$ level = .10 instead of .05 and $1 - \beta$ = .9 instead of .8), we would need at least **`r power_design2b$anova_n$n[3]` participants in each Learning instructions condition** ($N_{total}$ = `r power_design2b$anova_n$n[3]*2`) to have enough power ($1-\beta$ = `r power_design2b$anova_n$desired_power[3]/100`; calculated: `r round(power_design2b$anova_n$achieved_power[3]/100, digits=2)`) to find the interaction (estimated to be $\eta^2_p$ $\approx$ `r round(power_design2b$effect_sizes$partial_eta_squared[3], digits=3)`). 

This sample size provides enough power to find an EC effect of $\eta^2_p$ $\approx$ `r round(power_design2b$effect_sizes_emm$partial_eta_squared[2], digits=3)` in the **Valence** focus condition, as $1 - \beta$ = `r round(power_design2b$emm_n$achieved_power[2]/100, digits=2)` to find this effect is achieved with $n$ = `r power_design2b$emm_n$n[2]`. 

Regarding the EC effect in the **Age** focus condition, $n$ = `r power_design2b$emm_n$n[1]` `r ifelse(power_design2b$emm_n$n[1]==100, "(maximum n in the analysis)", "")` is `r ifelse(power_design2b$emm_n$achieved_power[1]<80, "not enough", "needed")` to detect it ($1-\beta$ = `r round(power_design2b$emm_n$achieved_power[1]/100, digits=2)`; $\eta^2_p$ $\approx$ `r round(power_design2b$effect_sizes_emm$partial_eta_squared[1], digits=3)`).

## Scenario 3: Scenario 2 but reduce the interaction effect even more

```{r setup_design3}
n = 50 #just to generate the anova design

r_age = cor(rat_age$mean_rating[rat_age$us_valence=="positive"],rat_age$mean_rating[rat_age$us_valence=="negative"]) #correlation between ratings on CS paired with US+ and US- in the age condition

r_val = cor(rat_val$mean_rating[rat_val$us_valence=="positive"], rat_val$mean_rating[rat_val$us_valence=="negative"]) #correlation between ratings on CS paired with US+ and US- in the valence condition

reduce_inter2 = .08 #changes EC effects by .16; reduces interaction by .32

r_mat = matrix(c(1,r_age,0,0
         ,r_age,1,0,0
         ,0,0,1,r_val
         ,0,0,r_val,1), nrow=4) #use values from the correlations calculated above 

design_exp3 = ANOVA_design(design = "2b*2w"
                          ,label_list = c("learn_instr"
                                          , "age", "valence"
                                          ,"us_val"
                                          ,"neg", "pos")
                           ,n = n
                           ,mu =c(rat_agg$mean[1]-reduce_inter2, rat_agg$mean[2]+reduce_inter2, rat_agg$mean[3]+reduce_inter2, rat_agg$mean[4]-reduce_inter2) 
                           ,sd = c(rat_agg$sd[1], rat_agg$sd[2], rat_agg$sd[3], rat_agg$sd[4])
                           ,r=r_mat
                           ,plot = TRUE)
```

```{r results_power3}
n_min = 10
n_max = 150
pwr = .8
a_lev = .05

#Find statistical power for a range of sample sizes
power_design3 = plot_power(design_exp3
                          ,min_n = n_min
                          ,max_n = n_max
                          ,desired_power = pwr*100
                          ,emm = TRUE
                          ,emm_comp = "us_val|learn_instr"
                          ,alpha_level = a_lev
                          ,plot = FALSE
                          ,verbose = FALSE) 

#power_design3$plot_emm

gt(power_design3$anova_n,caption = "Power analysis for the ANOVA -- Scenario 3")

gt(power_design3$effect_sizes,caption = "Effect sizes in the ANOVA -- Scenario 3")

gt(power_design3$emm_n,caption = "Power analysis for the post hoc analyses -- Scenario 3")

gt(power_design3$effect_sizes_emm,caption = "Effect sizes of the EC effects -- Scenario 3")
```

In **Scenario 3**, we would need at least **`r power_design3$anova_n$n[3]` participants in each Learning instructions condition** ($N_{total}$ = `r power_design3$anova_n$n[3]*2`) to have enough power ($1-\beta$ = `r power_design3$anova_n$desired_power[3]/100`; calculated: `r round(power_design3$anova_n$achieved_power[3]/100, digits=2)`) to find the interaction (estimated to be $\eta^2_p$ $\approx$ `r round(power_design3$effect_sizes$partial_eta_squared[3], digits=3)`). 

This sample size provides enough power to find an EC effect of $\eta^2_p$ $\approx$ `r round(power_design3$effect_sizes_emm$partial_eta_squared[2], digits=3)` in the **Valence** focus condition, as $1 - \beta$ = `r round(power_design3$emm_n$achieved_power[2]/100, digits=2)` to find this effect is achieved with $n$ = `r power_design3$emm_n$n[2]`. 

Regarding the EC effect in the **Age** focus condition, $n$ = `r power_design3$emm_n$n[1]` `r ifelse(power_design3$emm_n$n[1]==100, "(maximum n in the analysis)", "")` is `r ifelse(power_design3$emm_n$achieved_power[1]<80, "not enough", "needed")` to detect it ($1-\beta$ = `r round(power_design3$emm_n$achieved_power[1]/100, digits=2)`; $\eta^2_p$ $\approx$ `r round(power_design3$effect_sizes_emm$partial_eta_squared[1], digits=3)`).

## Scenario 4: Scenario 2 but r = .5

```{r setup_design4}
n = 50 #just to generate the anova design

r_age2 = .5
r_val2 = .5

reduce_inter3 = reduce_inter

r_mat2 = matrix(c(1,r_age2,0,0
         ,r_age2,1,0,0
         ,0,0,1,r_val2
         ,0,0,r_val2,1), nrow=4)

design_exp4 = ANOVA_design(design = "2b*2w"
                          ,label_list = c("learn_instr"
                                          , "age", "valence"
                                          ,"us_val"
                                          ,"neg", "pos")
                           ,n = n
                           ,mu = c(rat_agg$mean[1]-reduce_inter3, rat_agg$mean[2]+reduce_inter3, rat_agg$mean[3]+reduce_inter3, rat_agg$mean[4]-reduce_inter3) 
                           ,sd = c(rat_agg$sd[1], rat_agg$sd[2], rat_agg$sd[3], rat_agg$sd[4])
                           ,r=r_mat2
                           ,plot = TRUE)
```

```{r results_power4}
n_min = 10
n_max = 150
pwr = .8
a_lev = .05

#Find statistical power for a range of sample sizes
power_design4 = plot_power(design_exp4
                          ,min_n = n_min
                          ,max_n = n_max
                          ,desired_power = pwr*100
                          ,emm = TRUE
                          ,emm_comp = "learn_instr|us_val"
                          ,contrast_type = "pairwise"
                          ,alpha_level = a_lev 
                          ,plot = FALSE
                          ,verbose = FALSE) 

#power_design4$plot_emm

gt(power_design4$anova_n,caption = "Power analysis for the ANOVA -- Scenario 4")

gt(power_design4$effect_sizes,caption = "Effect sizes in the ANOVA -- Scenario 4")

gt(power_design4$emm_n,caption = "Power analysis for the post hoc analyses -- Scenario 4")

gt(power_design4$effect_sizes_emm,caption = "Effect sizes of the EC effects -- Scenario 4")
```

In **Scenario 4**, we would need at least **`r power_design4$anova_n$n[3]` participants in each Learning instructions condition** ($N_{total}$ = `r power_design4$anova_n$n[3]*2`) to have enough power ($1-\beta$ = `r power_design4$anova_n$desired_power[3]/100`; calculated: `r round(power_design4$anova_n$achieved_power[3]/100, digits=2)`) to find the interaction (estimated to be $\eta^2_p$ $\approx$ `r round(power_design4$effect_sizes$partial_eta_squared[3], digits=3)`). 

This sample size provides enough power to find an EC effect of $\eta^2_p$ $\approx$ `r round(power_design4$effect_sizes_emm$partial_eta_squared[2], digits=3)` in the **Valence** focus condition, as $1 - \beta$ = `r round(power_design4$emm_n$achieved_power[2]/100, digits=2)` to find this effect is achieved with $n$ = `r power_design4$emm_n$n[2]`. 

Regarding the EC effect in the **Age** focus condition, $n$ = `r power_design4$emm_n$n[1]` `r ifelse(power_design4$emm_n$n[1]==100, "(maximum n in the analysis)", "")` is `r ifelse(power_design4$emm_n$achieved_power[1]<80, "not enough", "needed")` to detect it ($1-\beta$ = `r round(power_design4$emm_n$achieved_power[1]/100, digits=2)`; $\eta^2_p$ $\approx$ `r round(power_design4$effect_sizes_emm$partial_eta_squared[1], digits=3)`).

# Additional analyses

## If the correlations between repeated measures are negative or Zero?

### Additional scenario 1: Scenario 1 with r = -0.5

```{r setup_add1}
n = 50 

r_age = -.5
r_val = -.5

r_mat = matrix(c(1,r_age,0,0
         ,r_age,1,0,0
         ,0,0,1,r_val
         ,0,0,r_val,1), nrow=4) #use values from the correlations calculated above 

design_exp_add1 = ANOVA_design(design = "2b*2w"
                          ,label_list = c("learn_instr"
                                          , "age", "valence"
                                          ,"us_val"
                                          ,"neg", "pos")
                           ,n = n
                           ,mu = c(rat_agg$mean[1], rat_agg$mean[2], rat_agg$mean[3], rat_agg$mean[4]) 
                           ,sd = c(rat_agg$sd[1], rat_agg$sd[2], rat_agg$sd[3], rat_agg$sd[4])
                           ,r=r_mat
                           ,plot = FALSE)
```

```{r results_power_add1}
n_min = 10
n_max = 150
pwr = .8
a_lev = .05

#Find statistical power for a range of sample sizes
power_design_add1 = plot_power(design_exp_add1
                          ,min_n = n_min
                          ,max_n = n_max
                          ,desired_power = pwr*100
                          ,emm = TRUE
                          ,emm_comp = "us_val|learn_instr"
                          ,alpha_level = a_lev
                          ,plot = FALSE
                          ,verbose = FALSE) 

#power_design_add1$plot_emm

gt(power_design_add1$anova_n,caption = "Power analysis for the ANOVA -- Additional scenario 1")

gt(power_design_add1$effect_sizes,caption = "Effect sizes in the ANOVA -- Additional scenario 1")

gt(power_design_add1$emm_n,caption = "Power analysis for the post hoc analyses -- Additional scenario 1")

gt(power_design_add1$effect_sizes_emm,caption = "Effect sizes of the EC effects -- Additional scenario 1")
```

### Additional scenario 2: Scenario 1 with r = 0

```{r setup_add2}
n = 50 

r_age = 0
r_val = 0

r_mat = matrix(c(1,r_age,0,0
         ,r_age,1,0,0
         ,0,0,1,r_val
         ,0,0,r_val,1), nrow=4) #use values from the correlations calculated above 

design_exp_add2 = ANOVA_design(design = "2b*2w"
                          ,label_list = c("learn_instr"
                                          , "age", "valence"
                                          ,"us_val"
                                          ,"neg", "pos")
                           ,n = n
                           ,mu = c(rat_agg$mean[1], rat_agg$mean[2], rat_agg$mean[3], rat_agg$mean[4]) 
                           ,sd = c(rat_agg$sd[1], rat_agg$sd[2], rat_agg$sd[3], rat_agg$sd[4])
                           ,r=r_mat
                           ,plot = FALSE)
```

```{r results_power_add2}
n_min = 10
n_max = 150
pwr = .8
a_lev = .05

#Find statistical power for a range of sample sizes
power_design_add2 = plot_power(design_exp_add2
                          ,min_n = n_min
                          ,max_n = n_max
                          ,desired_power = pwr*100
                          ,emm = TRUE
                          ,emm_comp = "us_val|learn_instr"
                          ,alpha_level = a_lev
                          ,plot = FALSE
                          ,verbose = FALSE) 

#power_design_add1$plot_emm

gt(power_design_add2$anova_n,caption = "Power analysis for the ANOVA -- Additional scenario 2")

gt(power_design_add2$effect_sizes,caption = "Effect sizes in the ANOVA -- Additional scenario 2")

gt(power_design_add2$emm_n,caption = "Power analysis for the post hoc analyses -- Additional scenario 2")

gt(power_design_add2$effect_sizes_emm,caption = "Effect sizes of the EC effects -- Additional scenario 2")
```