---
title : "Experiment 3"
author: "Karoline Bading & Marius Barth"
date  : "`r Sys.Date()`"
abstract: |
  Beware, this is a test script!
output:
  bookdown::html_document2:
    toc: yes
    toc_float: yes
    code_folding: hide
editor_options: 
  chunk_output_type: console
---



```{r setup, include = FALSE}
library(afex)
library(emmeans)
library(ggeffects)
library(papaja)
library(memoise)
library(HMMTreeC)
library(TreeBUGS)
library(lavaan)

project_root <- rprojroot::find_rstudio_root_file()
study_folder <- file.path(project_root, "studies", "wsw3-main")

set_sum_contrasts()

knitr::opts_chunk$set(
  echo = TRUE
  , message = FALSE
)
knitr::opts_knit$set(
  global.par = TRUE
)

source(file.path(project_root, "R", "mptinr_helper.R"))

file_cache <- cache_filesystem(path = file.path(project_root, ".rcache"), compress = "gzip", algo = "md5")
simpleMPT <- memoise(TreeBUGS::simpleMPT, cache = file_cache)
BayesFactorMPT <- memoise(TreeBUGS::BayesFactorMPT, cache = file_cache)
```

```{r global-par}
wesanderson::wes_palette("Zissou1", n = 3, type = "continuous") |>
  palette()
par(las = 1, cex = 1.2)
```


```{r exp3-read-data}
data_list <- readRDS(file.path(study_folder, "data", "data.rds"))
rating <- data_list$rating
mpt_data <- data_list$mpt_data
```


## Evaluative rating

```{r}
ftable(1 ~ task_focus , subset(rating, !duplicated(sid)))
# ftable(us_valence ~ sid + us_age, data = rating)
```


(ref:exp3-ratings-caption) Mean evaluative ratings for CSs as a function of *US valence* and *task focus*. Error bars represent 95% confidence intervals. Small points represent individual participants' mean evaluative ratings.

```{r exp3-ratings, fig.cap = "(ref:exp3-ratings-caption)"}
apa_beeplot(
  data = subset(rating, !is.na(us_valence))
  , id = "sid"
  , dv = "evaluative_rating"
  , factors = c("task_focus", "us_valence")
  , ylim = c(1, 8)
  , args_legend = list(x = "topleft")
)

anova_rating_model <- aov_4(
  formula = evaluative_rating ~ task_focus * (us_valence | sid)
  , data = subset(rating, !is.na(us_valence))
  , fun_aggregate = mean
)

marginal_means <- emmeans(
  anova_rating_model
  , specs = ~ us_valence + task_focus
)

ec_effects <- marginal_means |>
  contrast(
    list(
      "age focus"       = c(1, -1, 0, 0)
      , "valence focus" = c(0, 0, 1, -1)
    )
    , side = ">"
  ) |>
  apa_print()

anova_rating <- apa_print(anova_rating_model)

rating_wide <- data_list$rating_wide
```

Figure\ \@ref(fig:exp3-ratings) shows mean evaluative ratings.
We analyzed evaluative ratings using
a 2(*task focus*: valence vs. age) $\times$ 2 (*US valence*: positive vs. negative) mixed-factorial ANOVA
that revealed
a main effect *US valence*,
`r anova_rating$full_result$us_valence` that was qualified by
the two-way interaction *US valence* and *task focus*,
`r anova_rating$full_result$task_focus_us_valence`.
The main effect of *task focus* condition was not significant,
`r anova_rating$full_result$task_focus`.

EC effect in valence-focus condition:
`r ec_effects$full_result$valence_focus`.
EC effect in age-focus condition:
`r ec_effects$full_result$age_focus`.

```{r eval = FALSE}
lme_model <- lmer(
  formula = evaluative_rating ~ us_valence * task_focus + (us_valence | sid) + (1 | cs)
  , data = subset(rating, !is.na(us_valence)) 
)
summary(lme_model)
ggeffects::ggpredict(lme_model, terms = ~ task_focus + us_valence) |>
  plot()
```


## MPT model



```{r exp3-ml-models}
baseline_restrictions <- list(
  G_x1 = .125
  , G_x2 = .125
)

hypothesis_restrictions <- list(
  baseline = list()
  , D_equal = list(D = c("D_x1", "D_x2"))
  , C_equal = list(C = c("C_x1", "C_x2"))
  , d_equal = list(d = c("d_x1", "d_x2"))
  , a_equal = list(a = c("a_x1", "a_x2"))
  , b_equal = list(b = c("b_x1", "b_x2"))
  , D_age_0   = list(D_x1=0)
  , D_val_0   = list(D_x2=0)
  , C_age_0   = list(C_x1=0)
  , C_val_0   = list(C_x2=0)
  , d_age_0   = list(d_x1=0)
  , d_val_0   = list(d_x2=0)
  , b_age_5   = list(b_x1=0.5)
  , b_val_5   = list(b_x2=0.5)
  , a_age_5   = list(a_x1=0.5)
  , a_val_5   = list(a_x2=0.5)
)

models <- lapply(
  hypothesis_restrictions
  , FUN = function(x) {
    fit_mpt(
      data = mpt_data
      , model = file.path(study_folder, "WSW_exp3.eqn")
      , restrictions = c(baseline_restrictions, x)
    ) |>
      add_condition("Task focus" = c("age" = "_x1", "valence" = "_x2")) |>
      label_parameters(D = "$D$", C = "$C$", d = "d$", a = "$a$", b = "$b$")
  }
)
plot(models$baseline)

```

```{r}
models$baseline
```


```{r exp3-compare-ml-models}
compare(models)
```

## Bayesian MPT model

```{r exp3-simple-mpt}
baseline_restrictions <- list(
  "G_x1 = G_x2 = .125"
)

simple_mpt <- lapply(
  hypothesis_restrictions
  , FUN = function(x) {
    # return(lapply(c(baseline_restrictions, x), FUN = paste, collapse = " = "))
    simpleMPT(
      eqnfile = file.path(study_folder, "WSW_exp3.eqn")
      , data = colSums(mpt_data)
      , restrictions = legacy_restrictions(c(baseline_restrictions, x))
      , n.iter   = 2e5L
      , n.burnin = 1e5L
      , n.thin   = 2e1L
      , ppp      = 5e3L
      , n.chains = 4e0L
      , cores    = 4e0L
    )
  }
)
saveRDS(simple_mpt, file = file.path(study_folder, "model-objects", "simple-mpt.rds"))

BFs <- BayesFactorMPT(simple_mpt, cores = parallel::detectCores())
saveRDS(BFs, file = file.path(study_folder, "model-objects", "simple-mpt-BFs.rds"))

# Compare baseline model with other models
data.frame(
  model = paste0("Effect of task focus on ", gsub(names(BFs$posterior[-1L]), pattern = "_equal", replacement = ""))
  , logBF =  log(BFs$posterior[[1L]]) - log(BFs$posterior[-1L])
  , row.names = NULL
) |> within({
  strong_evidence <- ifelse(abs(logBF) > log(10), "yes", "no")
  BF_01 <- exp(-logBF)
  BF_10 <- exp( logBF)
}) |>
  apa_table()

```


## Latent-trait MPT model

```{r eval = FALSE}
model <- readRDS(file.path(study_folder, "model-objects", "trait-mpt.rds"))
summary(model)
```

## Mediation analysis

```{r  eval = FALSE}
individual_parameters <- as.data.frame(getParam(model, parameter = "theta"))
# individual_parameters$sid <- rownames(individual_parameters)

tt <- transformedParameters(model,
  list("p_us_identity = D*C", "p_us_valence = D*(1-C)*d", "p_us_id_or_val = D*C + D*(1-C)*d"),
  level = "individual"
) |>
  do.call(what = "rbind") |>
  colMeans()

branch_probabilities <- data.frame(
  p_us_valence = tt[grepl(names(tt), pattern = "p_us_valence")]
  , p_us_identity = tt[grepl(names(tt), pattern = "p_us_identity")]
  , p_us_id_or_val = tt[grepl(names(tt), pattern = "p_us_id_or_val")]
  , row.names = NULL
)

data <- cbind(data_list$mpt_data_hierarchical, individual_parameters, branch_probabilities)
data <- merge(data, as.data.frame(rating_wide), by = c("sid", "task_focus"))
```


```{r  eval = FALSE}
z <- function(x) {as.numeric(scale(x))}

data <- within(data, {
  Dd <- z(D) * z(d)
  DC <- z(D) * z(C)
  dC <- z(d) * z(C)
  DdC <- z(D) * z(d) * z(C)
})


mediation_syntax <-"
    # Regressions
    D ~ a_D*task_focus 
    C ~ a_C*task_focus
    d ~ a_d*task_focus
    # b ~ a_b*task_focus
    # a ~ a_a*task_focus
    Dd ~ a_Dd*task_focus
    DC ~ a_DC*task_focus
    dC ~ a_dC*task_focus
    DdC ~ a_DdC*task_focus
    # ec_effect ~ b_D*D + b_C*C + b_d*d + b_b*b + b_a*a + b_Dd*Dd + b_DC*DC + b_dC*dC + b_DdC*DdC + ade*task_focus
    ec_effect ~ b_D*D + b_C*C + b_d*d + b_Dd*Dd + b_DC*DC + b_dC*dC + b_DdC*DdC + ade*task_focus
    
    #Define New Parameters
    ab_D := a_D*b_D                                   #the product term is computed as a*b
    ab_C := a_C*b_C
    ab_d := a_d*b_d
    ab_Dd := a_Dd*b_Dd
    ab_DC := a_DC*b_DC
    ab_dC := a_dC*b_dC
    ab_DdC := a_DdC*b_DdC
    total_effect := ade + ab_D + ab_C + ab_d + ab_Dd + ab_DC + ab_dC + ab_DdC  #having defined ab, we can use this here.
"

sem(mediation_syntax, data = data) |>
  summary(standardized=TRUE)
```

```{r eval = FALSE}
ec_lm <- lm(formula = ec_effect ~ D * C * d, data = data)
ec_lm |>
  marginaleffects::plot_predictions(condition = c("d", "C"))

ec_lm |>
  marginaleffects::plot_predictions(condition = c("C", "d"))

emtrends(ec_lm, var = "d", specs = ~ C, at = list(C = c(0, .1, .2, .3, .8))) |>
  summary(infer = TRUE)

emtrends(ec_lm, var = "C", specs = ~ d, at = list(d = ppoints(n = 10))) |>
  summary(infer = TRUE)
```

```{r three-way-interaction,  eval = FALSE}
ec_lm |>
  marginaleffects::plot_predictions(condition = c("d", "C", "D"))

conditions <- list(
  d = seq(0, 1, length.out = 1e3)
  , C   = c(.2, .5, .8)
  , D = c(.2, .5, .8)
)

marginaleffects::plot_predictions(ec_lm, condition = conditions)
emtrends(ec_lm, var = "d", specs = ~ C + D, at = conditions[2:3]) |>
  summary(infer = TRUE)

```

```{r eval = FALSE}
par(mfrow = c(1, 3))
plot(d ~ C, data = data, pch = 16)
abline(a = 0, b = 1)
plot(d ~ D, data = data, pch = 16)
plot(C ~ D, data = data, pch = 16)
```



```{r eval = FALSE}
lm(formula = ec_effect ~ D * C * d + task_focus, data = data) |>
  ggeffect(terms = ~ d:D) |>
  plot(show_data = TRUE)
```




### with unconditional probabilities of US-valence and US-identity memory as predictors

```{r eval = FALSE}
# hist(data$p_us_identity)
# hist(data$p_us_valence)
data$ln_us_identity <- qnorm(data$p_us_identity)
data$ln_us_valence <- qnorm(data$p_us_valence)
data$ln_us_id_or_val <- log(data$p_us_id_or_val)
data$I_valence_memory <- data$ln_us_identity * data$ln_us_valence

mediation_syntax <-"
    # Regressions
    # D ~ a_D*task_focus 
    # C ~ a_C*task_focus
    # d ~ a_d*task_focus
    # b ~ a_b*task_focus
    # a ~ a_a*task_focus
    ln_us_valence  ~ a_us_valence*task_focus
    ln_us_identity ~ a_us_identity*task_focus
    # I_valence_memory ~ a_inter*task_focus
    ec_effect ~ b_us_valence*ln_us_valence + b_us_identity*ln_us_identity + ade*task_focus
    
    #Define New Parameters
    ab_us_valence := a_us_valence*b_us_valence
    ab_us_identity := a_us_identity*b_us_identity
    # ab_inter := a_inter*b_inter
    c := ade + ab_us_valence + ab_us_identity#  + ab_inter #having defined ab, we can use this here.
"

path_model <- sem(mediation_syntax, data = data)
path_model |>
  summary(standardized=TRUE)

# m1 <- lm(ln_us_identity ~ task_focus, data = data) |>
#   plot()
# m2 <- lm(ln_us_valence ~ task_focus, data = data) |>
#   plot()
lm(ec_effect ~ task_focus + ln_us_identity * ln_us_valence, data = data) |>
  ggpredict(terms = ~ ln_us_valence + ln_us_identity) |>
  plot(show_data = TRUE)

lm(ec_effect ~ task_focus + p_us_valence * p_us_identity, data = data) |>
  ggpredict(terms = ~ p_us_valence) |>
  plot(show_data = TRUE)
```


## TreeStan model

```{r}
model <- readRDS(file.path(study_folder, "model-objects", "treestan.rds"))
summary(model, pars = "beta")
```

