---
title: On the potential role of attitude-based guessing in the "Who-said-what?" model of evaluative conditioning 
author: Marius Barth
date: today
code-fold: true
editor_options: 
  chunk_output_type: console
---


Data-generating model

```{r}
#| label: setup

# Helper functions ----
simulate_participant <- function(
  D
  , C 
  , d
  , a
  , b
  , G = 1/4
  , n_cs
  , n_dist
  , attitudes = list(mean = 0, sd = 0)
  , attitude_based_guessing = FALSE
  , binary_guessing = FALSE
  , rating = c("8-point-likert", "binary-attitude-based", "binary-memory-dominant", "continuous")
  , sid
) {
  
  rating <- match.arg(rating, several.ok = FALSE)
  
  stimulus_type <- factor(
    c(rep("CS", n_cs),  rep("Distractor", n_dist))
  )

  cs_idx   <- stimulus_type == "CS"
  dist_idx <- stimulus_type == "Distractor"
  
  us_valence <- rep("new", n_cs + n_dist)
  us_valence[cs_idx] <- rep(c("positive", "negative"), length.out = n_cs)
  
  tau_attitude <- rep(NA_real_, n_cs + n_dist)
  tau_attitude[cs_idx] <- rnorm(n = n_cs, mean = attitudes$mean, sd = attitudes$sd) * 
    ifelse(us_valence[cs_idx] == "positive", 1, -1)
  tau_attitude[dist_idx] <- rnorm(n = n_dist, mean = 0, sd = attitudes$sd)
  
  
  discrimination <- rbinom(
    n = n_cs + n_dist
    , prob = D
    , size = 1
  )
  identity_memory <- valence_memory <- rep(0L, n_cs + n_dist)
  
  # CSs recognized as old...
  idx <- cs_idx & discrimination == 1L
  identity_memory[idx] <- rbinom(
    n = sum(idx)
    , prob = C
    , size = 1L
  )
  valence_memory[idx] <- rbinom(
    n = sum(idx)
    , prob = d
    , size = 1L
  )
  valence_memory[identity_memory == 1L] <- 1L
  
  # manifeste variablen
  old_new <- rep(NA, n_cs + n_dist)
  
  idx <- cs_idx & discrimination == 1L
  old_new[idx] <- "old"
  
  idx <- dist_idx & discrimination == 1L
  old_new[idx] <- "new"
  
  idx <- discrimination == 0L
  old_new[idx] <- sample(c("old", "new"), size = sum(idx), replace = TRUE, prob = c(b, 1-b)) # guessing old
  
  
  chosen_valence <- rep(NA, n_cs + n_dist)
  
  idx <- cs_idx & valence_memory == 1
  chosen_valence[idx] <- us_valence[idx]
  
  idx <- (cs_idx & valence_memory == 0 | dist_idx) & old_new == "old"

  if(attitude_based_guessing) {
    if(binary_guessing) {
      p_positive <- ifelse(tau_attitude[idx] > 0, 1, 0)
    } else {
      p_positive <- pnorm(tau_attitude[idx])
    }
  } else {
    p_positive <- a
  }
  
  chosen_valence[idx] <- as.character(factor(
    rbinom(n = sum(idx), size = 1, prob = p_positive)
    , levels = c(1, 0)
    , labels = c("positive", "negative")
  ))
  
  chosen_valence[old_new == "new"] <- ""
  
  correct_us <- rep(NA, n_cs + n_dist)
  
  # informed decision
  idx <- old_new == "old" & identity_memory == 1L
  correct_us[idx] <- "cor"
  
  # correct valence guessed
  idx <- old_new == "old" & identity_memory == 0 & chosen_valence == us_valence
  correct_us[idx] <- sample(c("cor", "incor"), prob = c(G, 1-G), size = sum(idx), replace = TRUE)
  
  # wrong valence guessed -> wrong US chosen
  idx <- identity_memory == 0L & old_new == "old" & chosen_valence != us_valence
  correct_us[idx] <- "incor"
  
  
  correct_us[old_new == "new"] <- "new"
  
  # evaluative ratings
  memory_influence <- rep(0, n_cs + n_dist)
  idx <- valence_memory == 1L
  
  memory_influence[idx] <- rnorm(n = sum(idx), mean = 1, sd = 1) * ifelse(us_valence[idx] == "positive", 1, -1)
  
  # evaluative_rating <- 1 + rbinom(n = n_cs + n_dist, size = 7, prob = pnorm(tau_attitude + memory_influence))
  if(rating == "binary-attitude-based") {
    evaluative_rating <- ifelse(tau_attitude > 0, "positive", "negative")
    evaluative_rating <- factor(evaluative_rating, levels = c("positive", "negative"))
  } else if (rating == "binary-memory-dominant") {
    evaluative_rating <- rep(NA, n_cs + n_dist)
    idx <- valence_memory == 1L
    evaluative_rating[ idx] <- as.character(us_valence[idx])
    evaluative_rating[!idx] <- ifelse(tau_attitude[!idx] > 0, "positive", "negative")
    evaluative_rating <- factor(evaluative_rating, levels = c("positive", "negative"))
  } else if(rating == "8-point-likert") {
    evaluative_rating <- round(pnorm(tau_attitude) * 7 + 1)
  } else {
    stop("rating not properly defined")
  }

  
  
  list2DF(list(
    stimulus_type = stimulus_type
    , us_valence = us_valence
    
    , tau_discrimination = discrimination
    , old_new = old_new
    
    , tau_valence_memory = valence_memory
    , chosen_valence = chosen_valence
    
    , tau_identity_memory = identity_memory
    , correct_us = correct_us
    , mpt_response = paste0(
      factor(us_valence, levels = c("positive", "negative", "new"), labels = c("PosUS", "NegUS", "New"))
      , factor(chosen_valence, levels = c("positive", "negative", ""), labels = c("pos", "neg", ""))
      , correct_us
    )
    , evaluative_rating = evaluative_rating
    , sid = rep(sid, length(stimulus_type))
  ))
}

simulate_experiment <- function(
  n_subjs
  , n_cs = 24
  , n_dist = 24
  , D = .6
  , C = .4
  , d = .2
  , G = 1/4
  , a = .5
  , b = .5
  , attitude_based_guessing = FALSE
  , attitudes = list(mean = 0, sd = 1)
  , binary_guessing = FALSE
  , rating = c("8-point-likert", "binary-attitude-based", "binary-memory-dominant", "continuous")
) {
  
  rating <- match.arg(rating, several.ok = FALSE)
  
  eta <- list(
    D   = D
    , C = C
    , d = d
    , G = G
    , a = a
    , b = b
  ) |>
    lapply(qnorm) |>
    list2DF()

  sigma <- list2DF(list(
    D   = .5
    , C = .5
    , d = .5
    , G = 0
    , a = .2
    , b = .2
  ))
  
  theta <- Map(
    n = rep(n_subjs, ncol(eta))
    , mean = eta
    , sd = sigma
    , f = rnorm
  ) |>
    lapply(pnorm) |>
    setNames(nm = names(sigma)) |>
    list2DF()
  
  Map(
    f = simulate_participant
    , D = theta$D
    , C = theta$C
    , d = theta$d
    , G = theta$G
    , a = theta$a
    , b = theta$b
    , n_cs = n_cs
    , n_dist = n_dist
    , sid = seq_along(theta$D)
    , attitude_based_guessing = attitude_based_guessing
    , attitudes = rep(list(attitudes), n_subjs)
    , binary_guessing = binary_guessing
    , rating = rating
  ) |>
    do.call(what = "rbind")
}


# global preferences ----
library(TreeStan)

model_file <- file.path(rprojroot::find_rstudio_root_file(), "model-equations", "wsw-6.eqn")
response_levels <- MPTinR::check.mpt(model_file)$eqn.order.categories

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , cache = TRUE
)
options(
  width = 250
  , digits = 3L
  , contrasts = c("contr.sum", "contr.poly")
)

```

Default simulation values

```{r}
x <- formals(simulate_experiment) |>
  lapply(as.character)

x$rating <- x$rating[[2]]

knitr::kable(
  list2DF(
    list(
      parameter = names(x)
      , value = x
    )
  )
)

```




## Simulation 3

Does consistent responding between assignment task and evaluative ratings inflate parameter $d$?
Thhis might be the case if participants have pre-existing attitudes that are not correlated with
experimentally assigned US valence.

```{r simulation-3}
set.seed(123L)

d <- simulate_experiment(
  n_subjs = 120
  , attitude_based_guessing = TRUE
  , attitudes = list(mean = 0, sd = 1)
)



mpt_data <- as.data.frame(unclass(table(d$sid, d$mpt_response)))

sim3_model <- TreeStan::fit_mpt(
  model = model_file
  , data = mpt_data
  , restrictions = list(G = 1/4)
  , parameterization = "latent_location"
)
summary(sim3_model)
```

### Results

Guessing in line with pre-existing attitudes does not affect model parameters.



# Simulation 4

Does consistent responding between assignment task and evaluative ratings inflate $d$
if responses are based on **conditioned attitudes**?
Conditioned attitudes are positively correlated with US valence, which might cause a problem.

```{r simulation-4}
set.seed(4L)

d <- simulate_experiment(
  n_subjs = 60
  , attitude_based_guessing = TRUE
  , attitudes = list(mean = 1, sd = 1)
)

d$mpt_response <- factor(d$mpt_response, levels = response_levels)
summary(d$evaluative_rating)

mpt_data <- as.data.frame(unclass(table(d$sid, d$mpt_response)))

sim4_model <- TreeStan::fit_mpt(
  model = model_file
  , data = mpt_data
  , restrictions = list(G = 1/4)
  , parameterization = "latent_location"
)
summary(sim4_model)

sim4_trial_model <- TreeStan::fit_mpt(
  model = model_file
  , data = d
  , response = "mpt_response"
  , formula = ~ 1 + (1 | sid)
  , restrictions = list(G = 1/4)
  , parameterization = "latent_location"
)
summary(sim4_trial_model)
```

### Results

Attitude-based guessing in line with *conditioned* (i.e., correlated)
attitudes inflates the *d* parameter (heavily).
Other model parameters seem to be unimpressed.



## Simulation 5

Given that attitude-based guessing may inflate parameter $d$,
can we reduce (or even eliminate) the bias by adjusting $a$ with evaluative ratings?
In this simulation, evaluations are given on an 8-point Likert scale (discretized true attitude).
Attitude-based guessing is modeled as probabilistic $P(\mathrm{"positive"}) = \Phi(\tau)$.

Here we use the trial-level MPT model where parameter $a$ (probability of guessing "positive")
is predicted by the evaluative rating while all other parameters are constant across evaluative-rating values.
This is the correction proposed by Klauer in his review.


```{r simulation-5}
set.seed(5L)

d <- simulate_experiment(
  n_subjs = 60
  , attitude_based_guessing = TRUE
  , attitudes = list(mean = 1, sd = 1)
)

# subject-level data
# response_levels <- MPTinR::check.mpt(file.path(rprojroot::find_rstudio_root_file(), "model-equations/wsw-6-wide.eqn"))$eqn.order.categories
# d$mpt_response_2 <- factor(paste0(d$evaluative_rating, "_", d$mpt_response), levels = response_levels)
# mpt_data <- as.data.frame(unclass(table(d$sid, d$mpt_response_2)))
# 
# sim5_model <- TreeStan::fit_mpt(
#   model = file.path(rprojroot::find_rstudio_root_file(), "model-equations/wsw-6-wide.eqn")
#   , data = mpt_data
#   , restrictions = list(G = 1/4)
#   , parameterization = "latent_location"
# )
# summary(sim5_model)

sim5_trial_model <- TreeStan::fit_mpt(
  model = model_file
  , data = d
  , response = "mpt_response"
  , formula = list(
    ~ 1 + (1 | sid)
    , a ~ evaluative_rating + (evaluative_rating | sid)
  )
  , restrictions = list(G = 1/4)
  , parameterization = "latent_location"
)

summary(sim5_trial_model)
```

### Results

If evaluative ratings are only affected by (conditioned) attitudes,
but US valence memory is not expressed in evaluative ratings,
the correction proposed by Klauer indeed very much aligns the estimation of $d$.
In fact, it is not yet clear whether there remains a small underestimation of $d$




# Simulation 6

This simulation repeats the last one, but measurement error/random variation is eliminated.
To this end, both evaluative ratings and attitude-based guessing are determined by true attitude $\tau$:
If $\tau > 0$, a "positive" response is given in the rating task; in the memory task, $a = 1$.
If $\tau \leq 0$, a "negative" response in given in the rating task, in the memory task, $a = 0$.

> What if both guess responses in the memory task and evaluative ratings flow directly from conditioned attitudes?

```{r simulation-6}
# TreeBUGS::withinSubjectEQN(
#   "model-equations/wsw-6.eqn"
#   , save = "model-equations/wsw-6-binary-ratings.eqn"
#   , constant = c("D", "C", "d", "G", "b")
#   , labels = c("positive", "negative")
# )
set.seed(6L)

d <- simulate_experiment(
  n_subjs = 60
  , attitude_based_guessing = TRUE
  , attitudes = list(mean = 1, sd = 1)
  , rating = "binary-attitude-based"
  , binary_guessing = TRUE
)

response_levels <- MPTinR::check.mpt(
  file.path(rprojroot::find_rstudio_root_file(), "model-equations/wsw-6-binary-ratings.eqn")
)$eqn.order.categories


# d$mpt_response_2 <- factor(paste0(d$evaluative_rating, "_", d$mpt_response), levels = response_levels)
# mpt_data <- as.data.frame(unclass(table(d$sid, d$mpt_response_2)))
# 
# sim6_model <- TreeStan::fit_mpt(
#   model = file.path(rprojroot::find_rstudio_root_file(), "model-equations/wsw-6-binary-ratings.eqn")
#   , data = mpt_data
#   , restrictions = list(G = 1/4) # , a = c("a_positive", "a_negative"))
#   , parameterization = "latent_location"
# )
# summary(sim6_model)

sim6_trial_model <- TreeStan::fit_mpt(
  model = model_file
  , data = d
  , response = "mpt_response"
  , formula = list(
    ~ 1 + (1 | sid)
    , a ~ evaluative_rating + (evaluative_rating | sid)
  )
  , restrictions = list(G = 1/4)
  , parameterization = "latent_location"
)

summary(sim6_trial_model)
```

### Results

Simulation\ 6 demonstrates that, if evaluative ratings deterministically come from attitudes,
the Klauer correction works---$d$ is estimated in an unbiased way.



# Simulation 7

Another simulation where ratings and guessing deterministically flow from latent states.
Compared with Simulation 6, evaluative ratings are not only determined by latent attitude, but
US valence memory is dominant (i.e., ratings are only affected by attitude if there is no pairing memory).

We again use Klauer's correction.
Because not only conditioned attitudes, but also US valence memory, affect evaluative ratings,
controlling $a$ might result in an over-correction (i.e., under-estimation of $d$).

```{r simulation-7}
set.seed(7L)

d <- simulate_experiment(
  n_subjs = 60
  , attitude_based_guessing = TRUE
  , attitudes = list(mean = 1, sd = 1)
  , rating = "binary-memory-dominant"
  , binary_guessing = TRUE
)

# subject-level model
# response_levels <- MPTinR::check.mpt(file.path(rprojroot::find_rstudio_root_file(), "model-equations/wsw-6-binary-ratings.eqn"))$eqn.order.categories
# d$mpt_response_2 <- factor(paste0(as.character(d$evaluative_rating), "_", d$mpt_response), levels = response_levels)
# mpt_data <- as.data.frame(unclass(table(d$sid, d$mpt_response_2)))
# 
# sim7_model <- TreeStan::fit_mpt(
#   model = file.path(rprojroot::find_rstudio_root_file(), "model-equations/wsw-6-binary-ratings.eqn")
#   , data = mpt_data
#   , restrictions = list(G = 1/4)
#   , parameterization = "latent_location"
# )
# summary(sim7_model)

# trial-level model
sim7_trial_model <- TreeStan::fit_mpt(
  model = model_file
  , data = d
  , response = "mpt_response"
  , formula = list(
    ~ 1 + (1 | sid)
    , a ~ evaluative_rating + (evaluative_rating | sid)
  )
  , restrictions = list(G = 1/4)
  , parameterization = "latent_location"
  # , joint_prior = TRUE
)

summary(sim7_trial_model)

```


### Results

If both US valence memory and conditioned attitudes affect evaluative ratings,
using evaluative ratings to predict $a$ results in severe under-estimation of $d$.



# Simulation 8

Compared with Simulation\ 7, 
attitudes are not conditioned, but pre-existing
(i.e., not correlated with experimentally manipulated US valence).

This scenario implements the idea that evaluative ratings are affected by attitudes,
but there is no such thing as conditioned attitudes.
We can therefore have a look if the over-correction persists under such a scenario.

```{r simulation-8}
set.seed(8L)

d <- simulate_experiment(
  n_subjs = 60
  , attitude_based_guessing = TRUE
  , attitudes = list(mean = 0, sd = 1)
  , rating = "binary-memory-dominant"
  , binary_guessing = TRUE
)


# Trial-level model
sim8_trial_model <- TreeStan::fit_mpt(
  model = model_file
  , data = d
  , response = "mpt_response"
  , formula = list(
    ~ 1 + (1 | sid)
    , a ~ evaluative_rating + (evaluative_rating | sid)
  )
  , restrictions = list(G = 1/4)
  , parameterization = "latent_location"
)

summary(sim8_trial_model)

```

### Results

US valence memory (parameter $d$) is severely under-estimated.

