---
title : 'Symmetrical vs. asymmetrical relational pairs: Pilot study'
author: "Marius Barth"
date  : "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc         : true
    toc_float   : true
---

```{r setup, include = FALSE}
library(TreeBUGS)
library(afex)
library(emmeans)
library(ggeffects)
library(papaja)
library(lme4)
library(car)

knitr::opts_chunk$set(
  echo = TRUE
  , cache = TRUE
)

study_folder <- file.path(
  rprojroot::find_rstudio_root_file()
  , "pilot_data_analyses"
  #, "Pilot (project seminar)"
)
```

```{r load-and-prepare-data}

#dat = readRDS("ind_par_rate.RDS")
dat <- readRDS("data_wsw_final_pilot.RDS")
length(unique(dat$url.srid))
names(dat)
table(dat$serious)
table(dat$pay_attention)
```

```{r}
# delta_group <- transformedParameters(
#   mpt_model
#   , transformedParameters = list(
#     "delta_p = p1_asymmetrical - p1_symmetrical"
#     , "delta_m = m1_asymmetrical - m1_symmetrical"
#     , "delta_g = g_asymmetrical - g_symmetrical"
#   )
#   , level = "group"
# )
# summary(delta_group, quantiles = c(0.05, .25, .50, .75, .95))
```






```{r}
# reshape individual parameter estimates
#individual_parameters <- as.data.frame(getParam(mpt_model, parameter = "theta"))
individual_parameters <- dat
parameter_names <- colnames(individual_parameters)[6:12]

#saveRDS(individual_parameters, file = "individual-parameters-1.rds")

parameters_long <- tidyr::pivot_longer(
  individual_parameters
  , cols = all_of(parameter_names)
  , names_to = c("parameter", "relation_pair")
  , names_sep = "_"
)
parameters_long$parameter <- gsub(parameters_long$parameter, pattern = "1", replacement = "")

parameters_wide <- tidyr::pivot_wider(
  parameters_long
  , id_cols = c("id", "relation_pair")
  , names_from = "parameter"
)

parameters_wider <- tidyr::pivot_wider(
  parameters_long
  , id_cols = "id"
  , names_from = c("relation_pair", "parameter")
)

rating_wide <- tidyr::pivot_wider(
  rating_agg
  , id_cols = c("id", "US_valence", "relation_type")
  , values_from = "rating"
  , names_from = "relation_pair"
  , names_prefix = "rating_"
)

rating_wide_2 <- tidyr::pivot_wider(
  rating_agg
  , id_cols = c("id", "US_valence", "relation_pair")
  , values_from = "rating"
  , names_from = "relation_type"
  , names_prefix = "rating_"
)

rating_wide_2$AC.diff <- abs(rating_wide_2$rating_assimilative) - abs(rating_wide_2$rating_contrastive)

rating_wide_3 <- tidyr::pivot_wider(
  rating_wide_2
  , id_cols = c("id", "US_valence")
  , values_from = "AC.diff"
  , names_from = "relation_pair"
  , names_prefix = ""
)

ratings_with_parameters <- merge(rating_agg, parameters_wide, by = c("id", "relation_pair"))
ratings_with_parameters_wider <- merge(rating_wide, parameters_wider)
ratings_with_parameters_wider_3 <- merge(rating_wide_3,parameters_wider)
```

```{r}
apa_barplot(data = rating_agg
            , id = "id"
            , dv = "rating"
            , factors = c("relation_pair","US_valence"))
```


## Linear models similar to Kukken et al.

Kukken et al.:

> The model included a fixed intercept, fixed effects for the *m* and *p* predictors, and fixed effects for the factors US valence and relation. Additionally, fixed effects for all interaction terms between US valence, relation, and the *m* parameter, and US valence, relation, and the *p* parameter were included. We followed Barr, Levy, Scheepers, and Tily's (2013) advice to use a maximal random-effects structure: The repeated measures nature of the data was accordingly modeled by including a per-participant random adjustment to the fixed intercept ("random intercept"), and per-participant random adjustments to the slopes of the factors US valence, relation, and their interaction. Furthermore, we included a random intercept for the CSs

This is what we implemented here with separate models for symmetrical vs. asymmetrical relational qualifiers.
<!-- Because we only had one CS per within-subjects condition, we dropped the random effects for CSs. -->
All continuous predictors were centered and we used effect coding for categorical predictors.


### Symmetrical relational qualifier

```{r}
contrasts(ratings_with_parameters$US_valence) <- "contr.sum"
contrasts(ratings_with_parameters$relation_type) <- "contr.sum"
contrasts(ratings_with_parameters$relation_pair) <- "contr.sum"
center <- function(x) { x - mean(x, na.rm = TRUE) }

linear_model <- lm(
  rating ~ center(p) * US_valence * relation_type + center(m) * US_valence * relation_type
  , data = subset(ratings_with_parameters, relation_pair == "symmetrical")
)

ratings_widest <- tidyr::pivot_wider(
  subset(ratings_with_parameters, relation_pair == "symmetrical")
  , id_cols = c("g", "m", "p")
  , names_from = c("relation_type", "US_valence")
  , values_from = "rating"
)

model <- lm(
  cbind(contrastive_positive, assimilative_positive, contrastive_negative, assimilative_negative) ~ center(p) + center(m)
  , data = ratings_widest
)

model_tests <- Anova(
  model
  , idata = data.frame(
    US_valence = factor(rep(c("positive", "negative"), each = 2))
    , relation_type = factor(rep(c("contrastive", "assimilative"), times = 2))
  )
  , idesign = ~ US_valence * relation_type
  , type = 3
  , multivariate = FALSE
)

# lme_model <- lmer(
#   formula = rating ~ center(p) * US_valence * relation_type + center(m) * US_valence * relation_type + (0 + relation_type:US_valence  | id)
#   , data = subset(ratings_with_parameters, relation_pair == "symmetrical")
# )

apa_out <- apa_print(model_tests)
apa_out$table$term[] <- gsub(apa_out$table$term, pattern = "Center|1", replacement = "")
apa_table(
  apa_out
  , caption = "Linear regression for the symmetrical relational qualifiers"
)





```

#### Simple-slopes analysis

```{r}
emtrends(
  linear_model
  , specs =  ~ US_valence + relation_type
  , var = "p"
) |> summary(infer = TRUE) |> knitr::kable()
```


```{r}
ggeffects::ggpredict(
  linear_model
  , terms = ~ p + US_valence + relation_type
) |> plot(rawdata = TRUE, jitter = 0)
```

```{r}
emtrends(
  linear_model
  , specs =  ~ US_valence + relation_type
  , var = "m"
) |> summary(infer = TRUE) |> knitr::kable()
```


```{r}
ggeffects::ggpredict(
  linear_model
  , terms = ~ m + US_valence + relation_type
) |> plot(rawdata = TRUE, jitter = 0)
```

### Asymmetrical relational qualifier

```{r}
linear_model <- lm(
  rating ~ center(p) * US_valence * relation_type + center(m) * US_valence * relation_type
  , subset(ratings_with_parameters, relation_pair == "asymmetrical")
)

ratings_widest <- tidyr::pivot_wider(
  subset(ratings_with_parameters, relation_pair == "asymmetrical")
  , id_cols = c("g", "m", "p")
  , names_from = c("relation_type", "US_valence")
  , values_from = "rating"
)

model <- lm(
  cbind(contrastive_positive, assimilative_positive, contrastive_negative, assimilative_negative) ~ center(p) + center(m)
  , data = ratings_widest
)

model_tests <- Anova(
  model
  , idata = data.frame(
    US_valence = factor(rep(c("positive", "negative"), each = 2))
    , relation_type = factor(rep(c("contrastive", "assimilative"), times = 2))
  )
  , idesign = ~ US_valence * relation_type
  , type = 3
  , multivariate = FALSE
)

apa_out <- apa_print(model_tests)
apa_out$table$term[] <- gsub(apa_out$table$term, pattern = "Center|1", replacement = "")
apa_table(
  apa_out
  , caption = "Linear regression for the asymmetrical relational qualifiers"
)
```


```{r}
ggeffects::ggpredict(
  linear_model
  , terms = ~ p + US_valence + relation_type
) |> plot(rawdata = TRUE, jitter = 0)
```

```{r}
ggeffects::ggpredict(
  linear_model
  , terms = ~ m + US_valence + relation_type
) |> plot(rawdata = TRUE, jitter = 0)
```

#### Simple-slopes analysis

<!-- todo -->

### Combined analysis of both symmetrical and asymmetrical relational qualifiers

```{r}
linear_model <- lm(
  rating ~ center(p) * US_valence * relation_type * relation_pair + center(m) * US_valence * relation_type * relation_pair 
  , ratings_with_parameters
)

ratings_widest_both <- tidyr::pivot_wider(
  ratings_with_parameters
  , id_cols = c("g", "m", "p", "id", "relation_pair")
  , names_from = c("relation_type", "US_valence")
  , values_from = "rating"
)
# Das hier sind die Daten:
ratings_widest_both <- tidyr::pivot_wider(
  ratings_widest_both
  , id_cols = "id"
  , names_from = "relation_pair"
  , values_from = c("g", "m", "p", "assimilative_negative", "assimilative_positive", "contrastive_negative", "contrastive_positive")
)
model <- lm(
  formula = cbind(
    assimilative_positive_symmetrical,
    assimilative_negative_symmetrical,
    contrastive_positive_symmetrical,
    contrastive_negative_symmetrical,
    assimilative_positive_asymmetrical,
    assimilative_negative_asymmetrical,
    contrastive_positive_asymmetrical,
    contrastive_negative_asymmetrical
  ) ~ center(m_symmetrical) + center(p_symmetrical) + center(m_asymmetrical) + center(p_asymmetrical)
  , data = ratings_widest_both
)
# Hier bin ich mir noch nicht sicher:
model_tests <- Anova(
  model
  , idata = data.frame(
    US_valence = factor(rep(c("positive", "negative"), times = 4, each = 1))
    , relation_type = factor(rep(c("assimilative", "contrastive"), times = 2, each = 2))
    , relation_pair= factor(rep(c("symmetrical", "asymmetrical"), times = 1, each = 4))
  )
  , idesign = ~ US_valence * relation_type * relation_pair
  , type = 3
  , multivariate = FALSE
)

apa_out <- apa_print(model_tests)
apa_out$table$term[] <- gsub(apa_out$table$term, pattern = "Center|1", replacement = "")
# apa_table(
#   apa_out
#   , caption = "Linear regression for both relational qualifiers combined"
# )
```


### Difference scores

For both model parameters and evaluative ratings, we calculated differences between symmetrical and asymmetrical conditions, i.e. $\Delta p = p_{\mathrm{asymmetrical}} - p_{\mathrm{symmetrical}}$.

```{r}
differences <- data.frame(
  id = ratings_widest_both$id
)

vars <- gsub(
  setdiff(colnames(ratings_widest_both), "id")
  , pattern = "_symmetrical|_asymmetrical$"
  , replacement = ""
) |> unique()

for (i in vars) {
  differences[[i]] <- ratings_widest_both[[paste0(i, "_asymmetrical")]] - ratings_widest_both[[paste0(i, "_symmetrical")]]
}

model <- lm(
  formula = cbind(
    assimilative_positive,
    assimilative_negative,
    contrastive_positive,
    contrastive_negative
  ) ~ center(m) + center(p)
  , data = differences
)
model_tests <- Anova(
  model
  , idata = data.frame(
    US_valence = factor(rep(c("positive", "negative"), times = 2, each = 1))
    , relation_type = factor(rep(c("assimilative", "contrastive"), times = 1, each = 2))
  )
  , idesign = ~ US_valence * relation_type
  , type = 3
  , multivariate = FALSE
)
apa_out <- apa_print(model_tests)
apa_out$table$term[] <- gsub(apa_out$table$term, pattern = "Center|1", replacement = "")
apa_table(apa_out)
```


# Analyse p-differences



```{r}
ratings_with_parameters_wider_3$p.diff <- ratings_with_parameters_wider_3$asymmetrical_p - ratings_with_parameters_wider_3$symmetrical_p
ratings_with_parameters_wider_3$m.diff <- ratings_with_parameters_wider_3$asymmetrical_m - ratings_with_parameters_wider_3$symmetrical_m
ratings_with_parameters_wider_3$rating.diff <- ratings_with_parameters_wider_3$asymmetrical - ratings_with_parameters_wider_3$symmetrical


CP <- subset(ratings_with_parameters_wider_3,US_valence =="positive")
cor.test(CP$p.diff,CP$rating.diff)
cor.test(CP$m.diff,CP$rating.diff)
CN <- subset(ratings_with_parameters_wider_3,US_valence =="negative")
cor.test(CN$p.diff,CN$rating.diff)
cor.test(CN$m.diff,CN$rating.diff)

linear_model <- lm(
   rating.diff ~ center(p.diff) * US_valence + center(m.diff) * US_valence
   , ratings_with_parameters_wider_3
)

ratings_widest <- tidyr::pivot_wider(
  ratings_with_parameters_wider_3
  , id_cols = c("m.diff", "p.diff")
  , names_from = c("US_valence")
  , values_from = "rating.diff"
)

model <- lm(
  cbind(positive, negative) ~ center(p.diff) + center(m.diff)
  , data = ratings_widest
)

model_tests <- Anova(
  model
  , idata = data.frame(
    US_valence = factor(rep(c("positive", "negative"), each = 1))
    #, relation_type = factor(rep(c("contrastive", "assimilative"), times = 2))
  )
  , idesign = ~ US_valence
  , type = 3
  , multivariate = FALSE
)

apa_out <- apa_print(model_tests)
apa_out$table$term[] <- gsub(apa_out$table$term, pattern = "Center|1", replacement = "")
apa_table(
  apa_out
  , caption = "Linear regression for the asymmetrical relational qualifiers"
)




# linear_model_2 <- lm(
#   rating.diff ~ center(p.diff) * US_valence + center(m.diff) * US_valence
#   , ratings_with_parameters_wider_3
# )
# 
# linear_model <- lm(
#   rating.diff ~ center(p.diff) * center(m.diff) * US_valence
#   , ratings_with_parameters_wider_3
# )
# 
# apa_out <- apa_print(linear_model)
# apa_out$table$term[] <- gsub(apa_out$table$term, pattern = "Center|1", replacement = "")
# apa_table(
#   apa_out
#   , caption = "Linear regression for both relational qualifiers combined"
# )



```

```{r}
ggeffects::ggpredict(
  linear_model_2
  , terms = ~ p.diff + US_valence
) |> plot(rawdata = TRUE, jitter = 0)
```

```{r}
ggeffects::ggpredict(
  linear_model_2
  , terms = ~ m.diff + US_valence
) |> plot(rawdata = TRUE, jitter = 0)
```


## Model allowing for interaction between parameters

### Symmetrical relational qualifier

```{r}
linear_model <- lm(
  rating ~ center(p) * center(m) * US_valence * relation_type
  , subset(ratings_with_parameters, relation_pair == "symmetrical")
)
apa_out <- apa_print(linear_model)
apa_out$table$term[] <- gsub(apa_out$table$term, pattern = "Center|1", replacement = "")
apa_table(
  apa_out
  , caption = "Linear regression for symmetrical relationcal qualifiers"
)
```

### Aymmetrical relational qualifier

```{r}
linear_model <- lm(
  rating ~ center(p) * center(m) * US_valence * relation_type
  , subset(ratings_with_parameters, relation_pair == "asymmetrical")
)
apa_out <- apa_print(linear_model)
apa_out$table$term[] <- gsub(apa_out$table$term, pattern = "Center|1", replacement = "")
apa_table(
  apa_out
  , caption = "Linear regression for asymmetrical relational qualifiers"
)
```


### Only asymmetrical relational qualifiers

```{r}
linear_model <- lm(
  rating ~ center(p) * center(m) * US_valence * relation_type
  , subset(ratings_with_parameters, relation_pair == "symmetrical")
)
apa_out <- apa_print(linear_model)
apa_out$table$term[] <- gsub(apa_out$table$term, pattern = "Center|1", replacement = "")
apa_table(
  apa_out
  , caption = "Linear regression for symmetrical relational qualifiers"
)
```


### Combined analysis of both symmetrical and asymmetrical relational qualifiers

```{r}
linear_model <- lm(
  rating ~ center(p) * center(m) * US_valence * relation_type * relation_pair
  , ratings_with_parameters
)
apa_out <- apa_print(linear_model)
apa_out$table$term[] <- gsub(apa_out$table$term, pattern = "Center|1", replacement = "")
apa_table(
  apa_out
  , caption = "Linear regression for both relational qualifiers combined"
)
```

```{r}
ggeffects::ggpredict(
  linear_model
  , terms = ~ p + US_valence + relation_type + relation_pair
) |> plot(rawdata = TRUE, jitter = 0)
```

```{r}
ggeffects::ggpredict(
  linear_model
  , terms = ~ m + US_valence + relation_type + relation_pair
) |> plot(rawdata = TRUE, jitter = 0)
```


## Supplementary analyses

### Would it help to increase the number of CSs per within-Ss condition?

To explore whether increasing the number of CSs in each within-Ss condition would help,
we doubled our data and pretended that we observed two evaluative ratings per within-Ss condition
(we also added some random noise to have a residual-variance component).

```{r fake-data}
set.seed(42L)
fake_data <- rbind(ratings_with_parameters, ratings_with_parameters)
fake_data$rating <- fake_data$rating + rnorm(nrow(fake_data))
contrasts(fake_data$US_valence) <- "contr.sum"
contrasts(fake_data$relation_type) <- "contr.sum"
contrasts(fake_data$relation_pair) <- "contr.sum"
```


```{r}
lme_symmetrical <- lmer(
  formula = rating ~ center(p) * US_valence * relation_type + center(m) * US_valence * relation_type + (relation_type * US_valence  | id)
  , data = subset(fake_data, relation_pair == "symmetrical")
)
summary(lme_symmetrical)
```

```{r}
lme_asymmetrical <- lmer(
  formula = rating ~ (center(p) + center(m)) * US_valence * relation_type + (relation_type * US_valence  | id)
  , data = subset(fake_data, relation_pair == "asymmetrical")
)
summary(lme_asymmetrical)
```

```{r}
lme_joint <- lmer(
  formula = rating ~ (center(p) + center(m)) * US_valence * relation_type * relation_pair + (relation_type * US_valence * relation_pair  | id)
  , data = fake_data
  , control = lmerControl(
    
  )
)
summary(lme_joint)
```

Model seems to converge and to be estimable. Hooray!
<!--
todo:
  (\Delta_eval_rating) ~ \Delta_p | symmetrisch - asymmetrisch
-->