---
title: "Pilot 2 for Experiment 3"
author: "Karoline Bading & Marius Barth"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    code_folding: hide
  pdf_document:
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyr)

library(afex)
library(emmeans)
library(ggeffects)
library(papaja)
library(memoise)
library(TreeBUGS)
library(lavaan)

set_sum_contrasts()

knitr::opts_chunk$set(
  echo = TRUE
  , message = FALSE
  , comment = NA
)

source(file.path(rprojroot::find_rstudio_root_file(), "R", "mptinr_helper.R"))
```


```{r}
study_folder <- file.path(rprojroot::find_rstudio_root_file(), "studies", "wsw3-p2")

data_list <- readRDS(file.path(study_folder, "data", "data.rds"))

rating <- data_list$rating
mpt_data <- data_list$mpt_data
mpt_data_hierarchical <- data_list$mpt_data_hierarchical
mpt_aggregated <- colSums(mpt_data)
```


## Evaluative rating

```{r}
apa_beeplot(
  data = subset(rating, !is.na(us_valence))
  , id = "sid"
  , dv = "evaluative_rating"
  , factors = c("task_focus", "us_valence")
  , ylim = c(1, 8)
  , args_legend = list(x = "topleft")
)

anova_rating_model <- aov_4(
  formula = evaluative_rating ~ task_focus * (us_valence | sid)
  , data = subset(rating, !is.na(us_valence))
  , fun_aggregate = mean
)

marginal_means <- emmeans(
  anova_rating_model
  , specs = ~ us_valence + task_focus
) |>
  contrast(
    list(
      "EC effect, age-focus condition"       = c(1, -1, 0, 0)
      , "EC effect, valence-focus condition" = c(0, 0, 1, -1)
    )
    , side = ">"
  )

anova_rating <- apa_print(anova_rating_model)

rating_wide <- data_list$rating_wide
marginal_means
```

main effect *task focus*,
`r anova_rating$full_result$task_focus`,
main effect *US valence*,
`r anova_rating$full_result$us_valence`,
two-way interaction *task focus* with *US valence*,
`r anova_rating$full_result$task_focus_us_valence`.

```{r eval = FALSE}
lme_model <- lmer(
  formula = evaluative_rating ~ us_valence * task_focus + (us_valence | sid) + (1 | cs)
  , data = subset(rating, !is.na(us_valence)) 
)
summary(lme_model)
ggeffects::ggpredict(lme_model, terms = ~ task_focus + us_valence) |>
  plot()
```



## MPT model

```{r}
restrictions_list <- list(
  x  = list("G_x1=.125","G_x2=.125")
  , Dageval = list("G_x1=.125","G_x2=.125", "D_x1=D_x2")
  , Cageval = list("G_x1=.125","G_x2=.125", "C_x1=C_x2")
  , dageval = list("G_x1=.125","G_x2=.125", "d_x1=d_x2")
  , aageval = list("G_x1=.125","G_x2=.125", "a_x1=a_x2")
  , bageval = list("G_x1=.125","G_x2=.125", "b_x1=b_x2")
  , Dage0   = list("G_x1=.125","G_x2=.125", "D_x1=0")
  , Dval0   = list("G_x1=.125","G_x2=.125", "D_x2=0")
  , Cage0   = list("G_x1=.125","G_x2=.125", "C_x1=0")
  , Cval0   = list("G_x1=.125","G_x2=.125", "C_x2=0")
  , dage0   = list("G_x1=.125","G_x2=.125", "d_x1=0")
  , dval0   = list("G_x1=.125","G_x2=.125", "d_x2=0")
  , bage5   = list("G_x1=.125","G_x2=.125", "b_x1=0.5")
  , bval5   = list("G_x1=.125","G_x2=.125", "b_x2=0.5")
  , aage5   = list("G_x1=.125","G_x2=.125", "a_x1=0.5")
  , aval5   = list("G_x1=.125","G_x2=.125", "a_x2=0.5")
)

models <- lapply(
  restrictions_list
  , FUN = function(x) {
    fit.mpt(
      data = mpt_aggregated
      , model.filename = file.path(study_folder, "WSW_exp3.eqn")
      , restrictions.filename = x
    )
  }
)

baseline_restrictions <- list(
  G_x1 = .125
  , G_x2 = .125
)

hypothesis_restrictions <- list(
  baseline = list()
  , D_equal = list(D = c("D_x1", "D_x2"))
  , C_equal = list(C = c("C_x1", "C_x2"))
  , d_equal = list(d = c("d_x1", "d_x2"))
  , a_equal = list(a = c("a_x1", "a_x2"))
  , b_equal = list(b = c("b_x1", "b_x2"))
  # , Dage0   = list("D_x1=0")
  # , Dval0   = list("D_x2=0")
  # , Cage0   = list("C_x1=0")
  # , Cval0   = list("C_x2=0")
  # , dage0   = list("d_x1=0")
  # , dval0   = list("d_x2=0")
  # , bage5   = list("b_x1=0.5")
  # , bval5   = list("b_x2=0.5")
  # , aage5   = list("a_x1=0.5")
  # , aval5   = list("a_x2=0.5")
)
library(HMMTreeC)
models <- lapply(
  hypothesis_restrictions
  , FUN = function(x) {
    fit_mpt(
      data = mpt_data
      , model = file.path(study_folder, "WSW_exp3.eqn")
      , restrictions = c(baseline_restrictions, x)
    )
  }
)
wesanderson::wes_palette("Zissou1", n = 3, type = "c") |> 
  palette()

plot(
  models$baseline
  , factors = list("Task focus" = c("age" = "_x1", "valence" = "_x2"))
  , parameters = c(
    D   = expression(italic(D))
    , C = expression(italic(C))
    , d = expression(italic(d))
    , a = expression(italic(a))
    , b = expression(italic(b))
  )
)

```

```{r}
model_comparisons <- compare(models)
model_comparisons
```

## Bayesian MPT model

```{r simple-mpt}
file_cache <- cache_filesystem(path = file.path(rprojroot::find_rstudio_root_file(), ".rcache"), compress = "gzip")
simpleMPT <- memoise(TreeBUGS::simpleMPT, cache = file_cache)
BayesFactorMPT <- memoise(TreeBUGS::BayesFactorMPT, cache = file_cache)

simple_mpt <- restrictions_list |>
  lapply(function(x) {
    simpleMPT(
      eqnfile = file.path(study_folder, "WSW_exp3.eqn")
      , data = colSums(mpt_data)
      , restrictions = x
      , n.iter   = 4e5L
      , n.burnin = 2e5L
      , n.thin   = 2e1L
      , ppp      = 2e3L
      , n.chains = 4e0L
      , cores    = 4e0L
    )
})
BFs <- BayesFactorMPT(simple_mpt, cores = 4L)
# Compare baseline model with other models
BFs$posterior[[1L]] / BFs$posterior[-1L]
log(BFs$posterior[[1L]]) - log(BFs$posterior[-1L])

```


## Hierarchical MPT model

```{r}
mpt_data_hierarchical <-merge(mpt_data_hierarchical, rating_wide, sort = FALSE) |>
  within({
    ec_valence_focus <- (task_focus == "valence") * ec_effect
    ec_age_focus <- (task_focus == "age") * ec_effect
  })
```

```{r}
model <- readRDS(file.path(study_folder, "model-objects", "trait-mpt.rds"))
```

```{r}
par(mfrow = c(1, 3))
plotFit(model)
plotFreq(model)
plotParam(model)
```

```{r}
summary(model)
```


## Hierarchical MPT model w/ EC effects as predictors

```{r}
# Estimate parameters
mpt_model_ec_predictor <-  readRDS(file.path(study_folder, "model-objects", "trait-mpt-with-ec.rds"))
```

```{r}
summary(mpt_model_ec_predictor)
```










## Mediation

```{r}
individual_parameters <- as.data.frame(getParam(mpt_model_ec_predictor, parameter = "theta"))
parameter_names <- colnames(individual_parameters)

tt <- transformedParameters(mpt_model_ec_predictor,
  list("p_us_identity = D*C", "p_us_valence = D*(1-C)*d"),
  level = "individual"
) |>
  do.call(what = "rbind") |>
  colMeans()

branch_probabilities <- data.frame(
  p_us_valence = tt[grepl(names(tt), pattern = "p_us_valence")]
  , p_us_identity = tt[grepl(names(tt), pattern = "p_us_identity")]
  , row.names = NULL
)

data <- cbind(mpt_data_hierarchical, individual_parameters, branch_probabilities)
```


```{r}
z <- function(x) {as.numeric(scale(x))}

data <- within(data, {
  Dd <- z(D) * z(d)
  DC <- z(D) * z(C)
  dC <- z(d) * z(C)
  DdC <- z(D) * z(d) * z(C)
})


mediation_syntax <-"
    # Regressions
    D ~ a_D*task_focus 
    C ~ a_C*task_focus
    d ~ a_d*task_focus
    # b ~ a_b*task_focus
    # a ~ a_a*task_focus
    Dd ~ a_Dd*task_focus
    DC ~ a_DC*task_focus
    dC ~ a_dC*task_focus
    DdC ~ a_DdC*task_focus
    # ec_effect ~ b_D*D + b_C*C + b_d*d + b_b*b + b_a*a + b_Dd*Dd + b_DC*DC + b_dC*dC + b_DdC*DdC + ade*task_focus
    ec_effect ~ b_D*D + b_C*C + b_d*d + b_Dd*Dd + b_DC*DC + b_dC*dC + b_DdC*DdC + ade*task_focus
    
    #Define New Parameters
    ab_D := a_D*b_D                                   #the product term is computed as a*b
    ab_C := a_C*b_C
    ab_d := a_d*b_d
    ab_Dd := a_Dd*b_Dd
    ab_DC := a_DC*b_DC
    ab_dC := a_dC*b_dC
    ab_DdC := a_DdC*b_DdC
    total_effect := ade + ab_D + ab_C + ab_d + ab_Dd + ab_DC + ab_dC + ab_DdC  #having defined ab, we can use this here.
"

sem(mediation_syntax, data = data) |>
  summary(standardized=TRUE)
```

```{r}
ec_lm <- lm(formula = ec_effect ~ D * C * d, data = data)
ec_lm |>
  marginaleffects::plot_predictions(condition = c("d", "C"))

ec_lm |>
  marginaleffects::plot_predictions(condition = c("C", "d"))

emtrends(ec_lm, var = "d", specs = ~ C, at = list(C = c(0, .1, .2, .3, .8))) |>
  summary(infer = TRUE)

emtrends(ec_lm, var = "C", specs = ~ d, at = list(d = ppoints(n = 10))) |>
  summary(infer = TRUE)
```

```{r}
lm(formula = ec_effect ~ D * C * d + task_focus, data = data) |>
  ggeffect(terms = ~ d:D) |>
  plot(show_data = TRUE)
```


How could we explain this interaction pattern?

- $d$ is a proxy for participants' eagerness to encode valence, so participants with small $d$ but high $C$ maybe did not encode valence (for both C and not-C items).
- 



### with unconditional probabilities of US-valence and US-identity memory as predictors

```{r}
# hist(data$p_us_identity)
# hist(data$p_us_valence)
data$ln_us_identity <- qnorm(data$p_us_identity)
data$ln_us_valence <- qnorm(data$p_us_valence)
data$I_valence_memory <- data$ln_us_identity * data$ln_us_valence

mediation_syntax <-"
    # Regressions
    # D ~ a_D*task_focus 
    # C ~ a_C*task_focus
    # d ~ a_d*task_focus
    # b ~ a_b*task_focus
    # a ~ a_a*task_focus
    ln_us_valence  ~ a_us_valence*task_focus
    ln_us_identity ~ a_us_identity*task_focus
    I_valence_memory ~ a_inter*task_focus
    ec_effect ~ b_us_valence*ln_us_valence + b_us_identity*ln_us_identity + b_inter*I_valence_memory + ade*task_focus
    
    #Define New Parameters
    ab_us_valence := a_us_valence*b_us_valence
    ab_us_identity := a_us_identity*b_us_identity
    ab_inter := a_inter*b_inter
    c := ade + ab_us_valence + ab_us_identity + ab_inter #having defined ab, we can use this here.
"

path_model <- sem(mediation_syntax, data = data)
path_model |>
  summary(standardized=TRUE)

# m1 <- lm(ln_us_identity ~ task_focus, data = data) |>
#   plot()
# m2 <- lm(ln_us_valence ~ task_focus, data = data) |>
#   plot()
lm(ec_effect ~ task_focus + ln_us_identity * ln_us_valence, data = data) |>
  ggpredict(terms = ~ ln_us_valence + ln_us_identity) |>
  plot(show_data = TRUE)
```

